---
title: "Pine Environment Summary"
author: "Katie Lotterhos"
date: "April 24, 2015"
output: html_document
---

This is a summary of the environmental data that was used for the bayenv2 analysis for the pine data.


```{r, echo=FALSE}
#install.packages("factoextra")
library(factoextra)
```

setwd("../enviPC")
### Principal components of the environmental variables used in this study


```{r}
env <- read.table("var_out_pine_all_COMBINED.table.contig_flt10.bayenv.envi2")
#head(env)

in2 <- t (env)
nam1 <- read.table("../enviPC/enviNamesAllAnalyses.txt")
dim(in2)
colnames(in2) <- nam1$V1
in3 <- prcomp (in2, scale. = T)
str(in3)
scores <- t (in3$x)

barplot(in3$sdev, ylab="Standard Deviation", names.arg = 1:22, xlab="PC axis")

# Eigenvalues
eig <- (in3$sdev)^2
# Variance in Percentage
variance <- eig*100/sum(eig)
# Cumulative variances
cumvar <- cumsum(variance)
perVar <- data.frame(eig = eig, variance = variance,
                     cumvariance = cumvar)
perVar
  # first 3 explain 80% of variance
```

### Correlation of variables onto Prin Comp
```{r}
pa <- fviz_pca_var(in3, )
#fviz_pca_var(in3, axes = c(2,3))
pb <- fviz_pca_var(in3, axes = c(1,3))

par(mfrow=c(2,1))
p1 = fviz_pca_var(in3, col.var="contrib")+
scale_color_gradient2(low="lightgrey", mid="grey", 
      high="blue", midpoint=4) + theme_minimal()

p2 <- fviz_pca_var(in3, col.var="contrib", axes=c(1,3))+
scale_color_gradient2(low="lightgrey", mid="grey", 
      high="blue", midpoint=4) + theme_minimal()

# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

multiplot(pa, pb, cols=1)

multiplot(p1, p2, cols=1)
  # colored by their average contribution to the variance in the two PC axes pictured

pdf("../results/PCloadings.pdf", width=5, height=8.5)
  multiplot(p1, p2, cols=1)
dev.off()

# with colors
# Helper function : 
# Correlation between variables and principal components
var_cor_func <- function(var.loadings, comp.sdev){
  var.loadings*comp.sdev
}

# Variable correlation/coordinates
  loadings <- in3$rotation
  apply(abs(loadings), 2, max)
    # maximum loading of a variable in first 3 PC axes is < 0.38
  loadings[,1:5]

sdev <- in3$sdev
sdev
var.coord <- var.cor <- t(apply(loadings, 1, var_cor_func, sdev))
head(var.coord[, 1:4])

var.cos2 <- var.coord^2
head(var.cos2[, 1:4])
comp.cos2 <- apply(var.cos2, 2, sum)
#The contribution of a variable to a given principal component is (in percentage) : (var.cos2 * 100) / (total cos2 of the component)
contrib <- function(var.cos2, comp.cos2){var.cos2*100/comp.cos2}
var.contrib <- t(apply(var.cos2,1, contrib, comp.cos2))
colSums(var.contrib)
var.contrib[, 1:4]

par(mfrow=c(1,1), mar=c(4,4,1,1))
plot(loadings[,1:4], var.contrib[, 1:4])
# maximum contribution to any one variable in PC1 is < 10%, in PC2 is < 15%, and in PC3 is < 15%
```

Contribution to variance of an individual environment $i$ to PC axis $j$ is:

$\frac{(l_{ij}\sigma_j)^2}{\sum_i (l_{ij}\sigma_j)^2}$

where l_i is the loading of environment i onto PC axis j, $\sigma_j$ is the standard deviation of PC axis j. 

#### Distribution of environments along PC axes
```{r}
loadings <- in3$rotation
dim(loadings)

```

#### Table and graphs of loadings for first 5 PC axes
```{r}
data.frame(as.character(nam1$V1), loadings[,1:5])
```

Note that the y-axis in the following graphs are meaningless, environments are just stacked for easy readability.

```{r}
  par(mfrow=c(3,1), mar=c(4,0,0,0))
  makePCplot <- function(pc){
    plot(loadings[,pc], (rank(loadings[,pc])), bty="l", yaxt="n", ylab="", xlab=paste("PC", pc), col=0, xlim=c(min(loadings[,pc])*1.1, max(loadings[,pc])*1.1))
    text(loadings[,pc], (rank(loadings[,pc])), as.character(nam1$V1))
  }
  for (i in 1:3){
  makePCplot(i)
  }
```

### Information about this R session
```{r}
sessionInfo()
```
