---
title: "Galaxy_Plots_on_AdaptreeEnvi_1R_CalcCorrs"
author: "Katie Lotterhos"
date: "1/4/2018"
output: html_document
---

setwd("/Users/katie/Desktop/CurrResearch/1-AdaptreeData/201509_PEGA/simulations/")

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if(!("ggplot2" %in% installed.packages())){install.packages("ggplot2")}
if(!("gstat" %in% installed.packages())){install.packages("gstat")}
if(!("sp" %in% installed.packages())){install.packages("sp")}   
if(!("maptools" %in% installed.packages())){install.packages("maptools")} 
if(!("ncf" %in% installed.packages())){install.packages("ncf")} 
if(!("igraph" %in% installed.packages())){install.packages("igraph")} 
if(!("ggraph" %in% installed.packages())){install.packages("ggraph")}
if(!("dplyr" %in% installed.packages())){install.packages("dplyr")}
if(!("extrafont" %in% installed.packages())){install.packages("extrafont")}
if(!("gRain" %in% installed.packages())){install.packages("gRain")}
if(!("bnlearn" %in% installed.packages())){install.packages("bnlearn")}

if(!("graph" %in% installed.packages())){
  source("https://bioconductor.org/biocLite.R")
  biocLite("graph", suppressUpdates = TRUE)   
  biocLite("Rgraphviz", suppressUpdates = TRUE)
  biocLite("RBGL", suppressUpdates = TRUE)
}

library(ggplot2)
library(gstat)
library(sp)
library(maptools)
library(ncf)

library(igraph)
library(ggraph)
library(graph)
library(Rgraphviz)
library(RBGL)
library(gRain)
library(bnlearn)
  require(dplyr)

  library(extrafont) 
  font_import() 
  loadfonts()
source("../analysis/myFunDist.R")
```


### Load data
```{r}
### envi
### This is the Adaptree environments overlaid onto our sampling design
env <- read.table("data/results_AdaptreeEnviFor_R90.txt", header=TRUE)
names(env)
names(env)[13] <- "MAP"
names(env)[14] <- "MSP"

headdir <- "data/LW2015data/"
```

### Function for performing co-association network analysis on sims
```{r}
EvaluateNetworkOnSims <- function(snps_file, headdir){
  print("++++++++++++++++++++++++++++++++++++++")
  print(snps_file)
snps <- read.table(paste(headdir,snps_file, sep=""))
env_snps <- read.table(paste(headdir,gsub(".lfmm", ".env", snps_file), sep=""))
results <- read.table(paste(headdir,gsub(".lfmm", "Bayenv2LFMMpca.Cpval", snps_file), sep=""), header=TRUE)
results <- results[results$UseSNP,]

  # snps is the allele frequency data
  # env_snps is the selection environment for the simulated selected loci
  # results are the resutls of association tests from L&W 2015
  env$SimEnv <- env_snps[seq(1,1800, by=20),1]
  
  #par(mfrow=c(1,1))
  #plot(env$Y_Pops, env$SimEnv)
  #  abline(lm(env$SimEnv~env$Y_Pops))
  #plot(env$Y_Pops, env$MAT)
  #  abline(lm(env$MAT~env$Y_Pops))
  #plot(env$X_Pops, env$MAT)
  #  abline(lm(env$MAT~env$X_Pops))

  #plot(env$X_Pops, env$SimEnv)

  #head(env)
  cor.with.sim <- cor(env$SimEnv, env[,c(2:3, 8:27)], method="spearman")
  print("Correlations between simulated environment and Adaptree environments")
  print(cor.with.sim)
  #The correlations between the adaptree environment and the simulated environment are quite low.

  ### Calculate rho
  ### Create a matrix to hold Spearman's rho and another matrix to hold $P$-values, where each row is a SNP and each column is an environmental variable
  ### Create environment matrix for individual data
  env2 <- matrix(NA, nrow(snps), 23) #hard coding
  k=0
  env_cols <- c(2:3, 8:28) #hard coding
  for (i in env_cols){
    k=k+1
    env2[,k] <- rep(env[,i], each=20)
  }
  dim(env2)
  colnames(env2) <- names(env)[env_cols]
  head(env2)

  ### Create matrix for results of associations between SNPs and Adaptree Environments
  ### and p-values for those associations
  TrueEnv.rho <- TrueEnv.P <- matrix(NA, ncol(snps), ncol(env2))
  colnames(TrueEnv.rho) <- names(env)[env_cols]
  colnames(TrueEnv.P) <- names(env)[env_cols]

  for (j in 1:ncol(env2)){
    print(colnames(env2)[j])
    for (i in 1:ncol(snps)){
        test <- cor.test(snps[,i], env2[,j], method="spearman", exact=FALSE)
        TrueEnv.rho[i,j] <- test$estimate
        TrueEnv.P[i,j] <- test$p.value
      }
  }

    ### Get indices of neutral and selected SNPs
  Neut = results$s_high==0
  print("Number neutral loci:")
  print(sum(Neut))

  Sel = results$s_high>0 
  print("Number selected loci:")
  print(sum(Sel))

  Neut_ind <- 1:9900
  (Sel_ind <- 9901:ncol(snps))

  ### How many are significant
  ### I determined which loci were "outliers" by the two criteria: 
  ### (1) p-values from Spearman's rho less than expected by Bonferroni correction. 
  ### (2) The first criteria plus a BF > 2 from the stucture-corrected association test with the truly selective environment

  ### False positives and true positives when truly selected environment is unknown
  (Bonferroni_P_N <- -log10(0.05/(10000*22)))
  results$Pos_P <- rowSums(-log10(TrueEnv.P[,1:22])>Bonferroni_P_N)>0
  
  results$Pos_BF <- results$log.bf>2
  
  FP_P <- sum(results$Pos_P[Neut]) #false positives based on p-value only
  FP_BF <- sum(results$Pos_BF[Neut])
  FP_P_BF <- sum(results$Pos_P[Neut] & results$Pos_BF[Neut])  # false positives based on p-value and BF
  
  print("Numer of false positives based on p-values")
  print(sum(FP_P))
  print("Numer of false positives based on BF")
  print(sum(FP_BF))
  print("Numer of false positives based on p-values and BF")
  print(sum(FP_P_BF))
  
  ### True Positivies
  TP_P <- sum(results$Pos_P[Sel]) #true positives based on p-value only
  TP_BF <- sum(results$Pos_BF[Sel])
  TP_P_BF <- sum(results$Pos_P[Sel] & results$Pos_BF[Sel])
  print("Numer of true positives based on p-values")
  print(sum(TP_P))
  print("Numer of true positives based on BF")
  print(sum(TP_BF))
  print("Numer of true positives based on p-values and BF")
  print(sum(TP_P_BF))
  
  # The false positive rate is really high in the simulated environment (not shown). Not sure why that is. It is true for the IBD sims, but not for the IM sims, which makes me think it has something to do with the scale of autocorrelation in the environment I simulated. Although, there are few or no false positives after correction for population structure at log10 BF > 2.

  # out <- results$Pos_P; filename=paste(snps_file, "Neut", sep="_")
ClusterAndNetworkEval <- function(out, disthreshold=0.1, filename, color=rep(adjustcolor("grey80", alpha=0.5), length(out))){  
  # out are the outlier set. May consider neutral only, or neutral and selected loci
  
  # Hierarchical clustering
  s_strength <- results$s_high[which(out)]
  forh <- TrueEnv.rho[which(out), 1:22]
  print(dim(forh))
  snpdist <- dist(abs(forh))
  h <- hclust(snpdist)
  snp_distmat <- as.matrix(snpdist)
  colnames(snp_distmat) <- which(out)
  rownames(snp_distmat) <- which(out)
  
  diag(snp_distmat)=NA
  snp_dists <- snp_distmat
  snp_dists[lower.tri(snp_dists)] <- NA
  #snp_dists[s_strength>0.01, s_strength>0.01] 
  
  #rowSums(snp_distmat<0.1, na.rm=TRUE)
  
  dists_neut_neut <- snp_dists[s_strength==0, s_strength==0]
  perc_dists_neut_neut_less_0.1 <- sum(dists_neut_neut<0.1, na.rm=TRUE)/sum(!(is.na(dists_neut_neut)))
  print(c("Percent of neutral-neutral pairwise distances < 0.1", perc_dists_neut_neut_less_0.1))
  #hist(dists_neut, breaks=seq(0,0.5, by=0.01), main=snps_file)
  
  dists_neut_sel <-  snp_dists[s_strength==0, s_strength>0]
  perc_dists_neut_sel_less_0.1 <- sum(dists_neut_sel<0.1, na.rm=TRUE)/sum(!(is.na(dists_neut_sel)))
  print(c("Percent of neutral-selected pairwise distances < 0.1", perc_dists_neut_sel_less_0.1))
  
  dists_sel_sel <- snp_dists[s_strength>0, s_strength>0]
  perc_dists_sel_sel_less_0.1 <- sum(dists_sel_sel<0.1, na.rm=TRUE)/sum(!(is.na(dists_sel_sel)))
  #hist(dists_all_s, breaks=seq(0,0.5, by=0.01), main=snps_file)
  print(c("Percent of selected-selected pairwise distances < 0.1", perc_dists_sel_sel_less_0.1))
  
  dists_s0.1_s0.1 <- snp_dists[s_strength==0.1, s_strength==0.1]
  perc_dists_s0.1_s0.1_less_0.1 <- sum(dists_s0.1_s0.1<0.1, na.rm=TRUE)/sum(!(is.na(dists_s0.1_s0.1)))
  #hist(dists_s0.1, breaks=seq(0,0.5, by=0.01), main=snps_file)
  print(c("Percent of all strong selected (s=0.1) loci pairwise distances < 0.1", perc_dists_s0.1_s0.1_less_0.1))
  
  d4_multi <- myFun(as.dist(snp_distmat))
     (nedges <- nrow(d4_multi %>%
    filter(abs(d4_multi$value) < disthreshold)))
  print("Number of edges")
  print(nedges)
    if(nedges > 30000){print("Too many edges"); break}
  
  ## Make graph
  graph_cors <- d4_multi %>%
    filter(d4_multi$value < disthreshold) %>% #filter to close distances
    graph_from_data_frame(directed = FALSE)
   V(graph_cors)$name
   
   ## Add nodes with no connections
  toadd <- !(rownames(snp_distmat) %in% V(graph_cors)$name)
    sum(toadd)
    graph_cors2 <- add_vertices(graph_cors, nv = sum(toadd, na.rm=TRUE), name= rownames(snp_distmat)[toadd])
    V(graph_cors2)$name
    sum(!(rownames(snp_distmat) %in% V(graph_cors2)$name))
    graph_cors <- graph_cors2
    
      #cluster membership
   mem=clusters(graph_cors)$membership
   
   # number edges for each node
   nodenumedge <- sapply(V(graph_cors)$name, function(x) length(E(graph_cors)[from(V(graph_cors)[x])]))
   
   m <- match(V(graph_cors2)$name, rownames(snp_distmat))

    ggraph(graph_cors) +
    geom_edge_fan2(aes(edge_alpha = 0.01)) +
    guides(edge_alpha = "none", edge_width = "none") +
    #scale_edge_colour_gradientn(limits = c(-1, 1), 
    #                            colors = c("firebrick2", "dodgerblue2")) +
    geom_node_point(size = 20, color=color[m]) +
    #geom_node_text(aes(label = V(graph_cors)$name, size=0.1)) +
    theme_graph() 
    
  ggsave(paste("results2/Network", filename,disthreshold, "dist.pdf", sep="_"), width=15, height=15, units="in")
  
   if(identical(names(nodenumedge), names(mem))){
    return(list(network = data.frame(gtcontig_pos=as.numeric(as.character(names(mem))),
                      subcluster=mem, 
                      numedge=nodenumedge
                      ),
                s_strength=s_strength,
                snp_dists=snp_dists,
                dists_neut_neut=dists_neut_neut,
                dists_sel_sel=dists_sel_sel,
                dists_neut_sel=dists_neut_sel,
                dists_s0.1_s0.1=dists_s0.1_s0.1,
                perc_dists_neut_neut_less_0.1=perc_dists_neut_neut_less_0.1,
                perc_dists_neut_sel_less_0.1=perc_dists_neut_sel_less_0.1,
                perc_dists_sel_sel_less_0.1=perc_dists_sel_sel_less_0.1,
                perc_dists_s0.1_s0.1_less_0.1=perc_dists_s0.1_s0.1_less_0.1
                ))
   }else{
     print("error: node names and edge names not identical")
   }
} # end function
   
# Neutral only cluster
clustNeut <- ClusterAndNetworkEval(results$Pos_P[Neut], filename=paste(snps_file, "Neut", sep="_"))
clustNeut$network$Type <- "Neut"
#str(clustNeut)

# FP and TP  cluster
  results$color <- NA
  results$color[results$Pos_P & Neut] <- adjustcolor("grey80", alpha=0.5)
  results$color[results$Pos_P & results$s_high==0.001] <- adjustcolor("palegreen", alpha=0.15)
  results$color[results$Pos_P & results$s_high==0.005] <- adjustcolor("seagreen2", alpha=0.5)
  results$color[results$Pos_P & results$s_high==0.01] <- adjustcolor("turquoise4", alpha=0.5)
  results$color[results$Pos_P & results$s_high==0.1] <- adjustcolor("navy", alpha=0.5)

clustAll <- ClusterAndNetworkEval(results$Pos_P, filename=paste(snps_file, "ALL", sep="_"), color=results$color[which(results$Pos_P)])

#str(clustAll)
clustAll$network <- clustAll$network[order(clustAll$network$gtcontig_pos),]
clustAll$network$s = results$s_high[clustAll$network$gtcontig_pos]



 png(file = paste("results2/SimNetworkBoxplot", snps_file,"_", 0.1,".pdf", sep=""), width=5, height=5, units="in", res=300)
  par(mar=c(4,4,0,0))
  boxplot(clustAll$network$numedge~clustAll$network$s, ylab="Number of edges", xlab="Selection strength")
  dev.off()

  return(list(snps_file=snps_file, results=results, cor.with.sim=cor.with.sim, TrueEnv.rho=TrueEnv.rho, TrueEnv.P=TrueEnv.P, 
              NumFP_P=sum(FP_P), NumFP_BF=sum(FP_BF), NumFP_P_BF=sum(FP_P_BF),
              NumTP_P=sum(TP_P), NumTP_BF=sum(TP_BF), NumTP_P_BF=sum(TP_P_BF), 
              clustNeut=clustNeut, clustAll=clustAll, snpdistmat = snp_distmat))

}# end function
```

### Loop through simulation files and call function

```{r}
headdir <- "data/LW2015data/"
x = list.files(headdir)
files <- x[grep("*.lfmm", x)]
out <- list()

for (i in 1:length(files)){
  snps_file <- files[i]
  out[[i]] <- EvaluateNetworkOnSims(snps_file, headdir)
}

save(out, file="results2/out.Rdata")
```

Next step: Analyze false positives and true positives from the simulations and their overall distribution of edges.

### Assess FP and TP rates
```{r}
load("results2/out.Rdata") #optional, can start here

str(out[[1]])
(file = as.character(unlist(lapply(out, function(x){x$snps_file}))))
(NumFP_P <- as.numeric(unlist(lapply(out, function(x){x$NumFP_P}))))
(NumFP_BF <- as.numeric(unlist(lapply(out, function(x){x$NumFP_BF}))))
(NumFP_P_BF <- as.numeric(unlist(lapply(out, function(x){x$NumFP_P_BF}))))
(NumTP_P <- as.numeric(unlist(lapply(out, function(x){x$NumTP_P}))))
(NumTP_BF <- as.numeric(unlist(lapply(out, function(x){x$NumTP_BF}))))
(NumTP_P_BF <- as.numeric(unlist(lapply(out, function(x){x$NumTP_P_BF}))))
myres1 <- data.frame(file, NumFP_P, NumFP_BF, NumFP_P_BF, NumTP_P, NumTP_BF, NumTP_P_BF)
myres1$demog <- c(rep("1R",3), rep("2R", 3), rep("IBD", 3), rep("IM", 3))
myres1

myres <- myres1[1:9,] # exclude IM
myres$demog <- factor(myres$demog, levels=c("IBD", "1R", "2R"), ordered=TRUE)
str(myres)

pdf("SimErrorRates.pdf", width=6, height=6)
  par(mfrow=c(2,2), mar=c(3,2.5,1,1), oma=c(3,3,3,0), las=1)
  boxplot(myres$NumFP_P/9900~ myres$demog, ylim=c(0,0.01), main="Bonferroni", ylab="False Positive Rate")
  text(0.5,0.01, "A")
  boxplot(myres$NumFP_P_BF/9900~ myres$demog, ylim=c(0,0.01), main="Bonferroni + BF")
    text(0.5,0.01, "B")
  boxplot(myres$NumTP_P/100~ myres$demog, ylim=c(0,1), ylab="True Positive Rate")
      text(0.5,1, "C")
  boxplot(myres$NumTP_P_BF/100~ myres$demog, ylim=c(0,1))
        text(0.5,1, "D")
  mtext("Demography", outer=TRUE, side=1)
  mtext("False positive rate", outer=TRUE, side=2, adj=0.9, las=0, line=1)
  mtext("True positive rate", outer=TRUE, side=2, adj=0.2, las=0, line=0)
dev.off()
```

### Visualize network complexity
```{r}
# NOTE HARD CODING FOR DEMOG
edges1RNeut <- unlist(lapply(1:3, function(x){out[[x]]$clustNeut$network$numedge}))
edges1RSel <- unlist(lapply(1:3, function(x){out[[x]]$clustAll$network$numedge}))
edges1RSel_s <- unlist(lapply(1:3, function(x){out[[x]]$clustAll$network$s}))

edges2RNeut <- unlist(lapply(4:6, function(x){out[[x]]$clustNeut$network$numedge}))
edges2RSel <- unlist(lapply(4:6, function(x){out[[x]]$clustAll$network$numedge}))
edges2RSel_s <- unlist(lapply(4:6, function(x){out[[x]]$clustAll$network$s}))

edgesIBDNeut <- unlist(lapply(7:9, function(x){out[[x]]$clustNeut$network$numedge}))
edgesIBDSel <- unlist(lapply(7:9, function(x){out[[x]]$clustAll$network$numedge}))
edgesIBDSel_s <- unlist(lapply(7:9, function(x){out[[x]]$clustAll$network$s}))

pdf("SelectionStrengthVs.Degree.pdf", width=8, height=3)
  par(mfrow=c(1,3), mar=c(3,3,3,1), oma=c(2,2,0,0))
  col=c(adjustcolor("grey80", alpha=0.5), adjustcolor("palegreen", alpha=0.15), adjustcolor("seagreen2", alpha=0.5), adjustcolor("turquoise4", alpha=0.5), adjustcolor("navy", alpha=0.5))
  boxplot(edgesIBDSel~edgesIBDSel_s, col=col, main="A) IBD")
  boxplot(edges1RSel~edges1RSel_s, col=col[c(1,3,4,5)], main="B) 1 refuge")
  boxplot(edges2RSel~edges2RSel_s, col=col[c(1,3,4,5)], main="C) 2 refugia")
  mtext("Strength of selection (0 = neutral)", 1, outer=TRUE)
  mtext("Degree in network", 2, outer=TRUE, las=0)
dev.off()
```

## Connectivity analysis
Percent of pairwise comparisons less than the distance threshold for different groups
```{r}
myres1$dists_neut_neut <- as.numeric(unlist(lapply(out, function(x){x$clustAll$perc_dists_neut_neut_less_0.1})))

myres1$dists_neut_sel <- as.numeric(unlist(lapply(out, function(x){x$clustAll$perc_dists_neut_sel_less_0.1})))

myres1$dists_sel_sel <- as.numeric(unlist(lapply(out, function(x){x$clustAll$perc_dists_sel_sel_less_0.1})))

myres1$dists_sel0.1_sel0.1 <- as.numeric(unlist(lapply(out, function(x){x$clustAll$perc_dists_s0.1_s0.1_less_0.1})))

names(myres1)

#note hard coding here for columns and names
distslong <- reshape(myres1[,8:12],direction="long", varying = names(myres1[9:12]), v.names="dist", timevar="Compare", times=c("Neut-Neut", "Neut-Sel", "Sel-Sel", "strongSel-\nstrongSel"))

## Proportion of pairwise distances
head(distslong)
pdf("Pairwisedistances.pdf", height=4, width=6)
  par(mfrow=c(1,3), mar=c(6,2,3,0), oma=c(3,3,0,1))
  boxplot(dist ~ Compare, data=distslong[distslong$demog=="IBD",], las=3, main="IBD", ylab="Proportion of pairwise distances less than 0.1", ylim=c(0,0.8))
  boxplot(dist ~ Compare, data=distslong[distslong$demog=="1R",], las=3, main="1R", ylim=c(0,0.8))
  boxplot(dist ~ Compare, data=distslong[distslong$demog=="2R",], las=3, main="2R", ylim=c(0,0.8))
  mtext("Proportion of pairwise distances less than 0.1", side=2,
        outer=TRUE, cex=0.8)
  mtext("Type of pairwise comparison", side=1, outer=TRUE)
dev.off()

tapply(myres1$dists_neut_neut, myres1$demog, mean)
```


```{r}
sessionInfo()
```
